An efficient composition of bidirectional programs by tupling and lazy updates

### 2020-02-20 paper/2.minbigul.tex
before renaming vars
insert newline to author names, spell checked

        \noindent $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} = \\
            \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$

@@

        \noindent $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
            \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$

@@

Also in the semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$, the return statements are used to express clearly the value of functions. Variables $s'$ and $v'$ wrapped in these returns are neccessary for checking the fi conditions.

@@

$\tab \textit{bfoldr} \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
$\tabs{3} \ cond_{sv} = \lambda (s_1,s_2).\lambda v.(s_1 = [\ ]), \ cond_s = \lambda (s_1,s_2).(s_1 = [\ ])$\\
$\tabs{3} \ bx_1 = RearrV \ g_1 \ g_2 \ bx_v$ where: \\
    $\tabs{4} g_1 = g_2^{-1} = \lambda [v].(v,[\ ]), bx_v = (Skip \ (\lambda \_.())) \times Replace$\\
$\tabs{3} \ bx_2 = RearrS \ f_1 \ f_2 \ bx_s$ where:\\
    $\tabs{4} f_1 = f_2^{-1} = \lambda ((x:xs),e).(x,(xs,e))$, 
$bx_s = ((Replace \times bfoldr \ bx) \circ bx)$

@@

$\tab foldr :: (a \to b \to b) \to b \to [a] \to b \\
\tabs{2} \ \ foldr \ f \ e \ [\ ] = e \\ 
\tabs{2} \ \ foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$

@@

Using $\textit{foldr}$, we can define other functions, for example:

$\tab reverse = foldr \ snoc \ [\ ]$

With $\textit{bfoldr}$, we can also write the bidirectional version $breverse$ as follows:

$\tab breverse = RearrS \ f_1 \ f_2 \ bx$ where:\\
    $\tabs{4} f_1 = f_2^{-1} = \lambda s \to (s, [\ ]), \tab bx = bfoldr \ bsnoc$

>>>>

        \noindent $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} = \\
            \tab (\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2})$

@@

        \noindent $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
            \tab (\getbx{bx_1}{s_1},\getbx{bx_2}{s_2})$

@@

Also in the semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$, the return statements are used to express clearly the value of functions. Variables $s'$ and $v'$ wrapped in these returns are necessary for checking the fi conditions.

@@

$\tab \textit{bfoldr} \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
$\tabs{3} \ cond_{sv} = \lambda (s_1,s_2).\lambda v.(s_1 = [\ ]), \ cond_s = \lambda (s_1,s_2).(s_1 = [\ ])$\\
$\tabs{3} \ bx_1 = RearrV \ g_1 \ g_2 \ bx_v$ where: \\
    $\tabs{4} g_1 = g_2^{-1} = \lambda [v].(v,[\ ]), bx_v = (Skip \ (\lambda \_.())) \times Replace$\\
$\tabs{3} \ bx_2 = RearrS \ f_1 \ f_2 \ bx_s$ where:\\
    $\tabs{4} f_1 = f_2^{-1} = \lambda ((x:xs),e).(x,(xs,e))$, 
$bx_s = ((Replace \times \textit{bfoldr} \ bx) \circ bx)$

@@

$\tab \textit{foldr} :: (a \to b \to b) \to b \to [a] \to b \\
\tabs{2} \ \ \textit{foldr} \ f \ e \ [\ ] = e \\ 
\tabs{2} \ \ \textit{foldr} \ f \ e \ (x:xs) = f \ x \ (\textit{foldr} \ f \ e \ xs)$

@@

Using $\textit{foldr}$, we can define other functions like $reverse = \textit{foldr} \ snoc \ [\ ]$. With $\textit{bfoldr}$, we can also write the bidirectional version $breverse$ as follows:

$\tab breverse = RearrS \ f_1 \ f_2 \ bx$ where: \\
$\tabs{4} f_1 = f_2^{-1} = \lambda s \to (s, [\ ]), \tab bx = \textit{bfoldr} \ bsnoc$

### 2020-02-20 paper/4.xpg.tex
insert newline to author names, spell checked

The problem for $cpg$ lies in redundant computations during the evaluation. To prevent such redundant computations from occuring, we introduce an extension named $kpg$. While $cpg$ evaluates values eagerly, $kpg$ does the opposite. Every value is evaluated lazily in a computation of $kpg$. The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, lks, lkv, s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $lks$ and $lkv$ are used for lazy evaluation of actual values. The output of this function is also a 6-tuple $(ks, kv, lks, lkv, s, v)$.\\
Suppose that we have a source $s_0$ and a view $v_0$. The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:

@@

\textcolor{red}{Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
  Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
  However, if we accumulate $lks$ and $lkv$, both might be evlauated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$. This includes the same computation. To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$. and $\kpginline{bx_2}$}
%It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
Second is in $\kpginline{Case}$ and not shown in the definition. We need to evaluate $lks' \ s'$ and $lkv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple. Such evaluations can be done lazily to make programs run faster. We use the above small optimizations in our implementation.

>>>>

The problem for $cpg$ lies in redundant computations during the evaluation. To prevent such redundant computations from occurring, we introduce an extension named $kpg$. While $cpg$ evaluates values eagerly, $kpg$ does the opposite. Every value is evaluated lazily in a computation of $kpg$. The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, lks, lkv, s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $lks$ and $lkv$ are used for lazy evaluation of actual values. The output of this function is also a 6-tuple $(ks, kv, lks, lkv, s, v)$.\\
Suppose that we have a source $s_0$ and a view $v_0$. The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:

@@

\textcolor{red}{Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
  Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
  However, if we accumulate $lks$ and $lkv$, both might be evaluated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$. This includes the same computation. To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$. and $\kpginline{bx_2}$}
%It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
Second is in $\kpginline{Case}$ and not shown in the definition. We need to evaluate $lks' \ s'$ and $lkv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple. Such evaluations can be done lazily to make programs run faster. We use the above small optimizations in our implementation.

### 2020-02-20 paper/main.tex
insert newline to author names, spell checked

\author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and \break
  {Bach Nguyen Trong\inst{1}\orcidID{0000-0002-6342-6020}} \and \break
  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and
  {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}

@@

%\begin{itemize}
%\item BiGUL 
%\item pg (pair of put and get)
%\item cpg (extention of pg. Keep eval information for reusing)
%\item kpg (extention of cpg. Lazy evaluation)
%\item xpg (combination of pg and kpg)
%\item Experiment
%\item Related work
%\item Conclusion and Future work
%\end{itemize}

>>>>

\author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and \break
  {Bach Nguyen Trong\inst{1}\orcidID{0000-0002-6342-6020}} \and \break
  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and \break
  {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}

@@

%\begin{itemize}
%\item BiGUL 
%\item pg (pair of put and get)
%\item cpg (extension of pg. Keep eval information for reusing)
%\item kpg (extension of cpg. Lazy evaluation)
%\item xpg (combination of pg and kpg)
%\item Experiment
%\item Related work
%\item Conclusion and Future work
%\end{itemize}
